Ordinal bhsdtr tutorial

### 2020-02-18 ordinal/Readme.Rmd
preprint links

In an SDT model we usually assume that the stimulus class affects
*only* the *d'* parameter, and that both criteria (gamma) and *d'* may
differ between the participants (or items, etc.). In bhsdtr *d'* is
represented as *log(d') = delta*, which forces it to be
non-negative. This is an important assumption of Signal Detection
Theory - if this assumption is violated, for example, because of the
response reversal in some participants or conditions, then an SDT
model is not valid and a non-trivial generalization of SDT should be
used instead (see the bhsdtr preprint for a more detailed explanation
and some references).

@@

We have fitted a hierarchical SDT model which respects the
non-negativity assumption, the ordering of the thresholds, accounts
for the log-normal distribution of the *d'* random effects, allows for
random effects in both the *d'* and the individual threshold
parameters while still respecting the SDT assumptions (thanks to an
order-preserving gamma link function and the delta log link function
which respects the non-negativity assumption), and accounts for the
correlations between the random effects. First such model was
described in the bhsdtr preprint. This model, however, does not assume
that the stimulus class affects only the *d'* parameter. We can now
again see if there is any evidence of non-selectivity:

@@

Among the other problems that need to be addressed some have to do
with non-linearity and aggregation. If you fit an SDT model, or any
other non-linear model, to data aggregated over grouping factors, such
as participants or items, because of non-linearity, which does not
play well with aggregation, some or all of the point and interval
estimates may be biased, perhaps severely so, and the inference may be
invalid. You can see how horribly biased the point *and* interval
estimates of every SDT parameter may become when you overly aggregate
your data in the bhsdtr preprint. By the way, I am now in the process
of examining the direction and magnitude of bias in SDT models fitted
to overly-aggregated data by using the excellent Confidence database
created by Doby Rahnev.

>>>>

In an SDT model we usually assume that the stimulus class affects
*only* the *d'* parameter, and that both criteria (gamma) and *d'* may
differ between the participants (or items, etc.). In bhsdtr *d'* is
represented as *log(d') = delta*, which forces it to be
non-negative. This is an important assumption of Signal Detection
Theory - if this assumption is violated, for example, because of the
response reversal in some participants or conditions, then an SDT
model is not valid and a non-trivial generalization of SDT should be
used instead (see the bhsdtr
[preprint](http://dx.doi.org/10.23668/psycharchives.2725) for a more
detailed explanation and some references).

@@

We have fitted a hierarchical SDT model which respects the
non-negativity assumption, the ordering of the thresholds, accounts
for the log-normal distribution of the *d'* random effects, allows for
random effects in both the *d'* and the individual threshold
parameters while still respecting the SDT assumptions (thanks to an
order-preserving gamma link function and the delta log link function
which respects the non-negativity assumption), and accounts for the
correlations between the random effects. First such model was
described in the bhsdtr
[preprint](http://dx.doi.org/10.23668/psycharchives.2725). This model,
however, does not assume that the stimulus class affects only the *d'*
parameter. We can now again see if there is any evidence of
non-selectivity:

@@

Among the other problems that need to be addressed some have to do
with non-linearity and aggregation. If you fit an SDT model, or any
other non-linear model, to data aggregated over grouping factors, such
as participants or items, because of non-linearity, which does not
play well with aggregation, some or all of the point and interval
estimates may be biased, perhaps severely so, and the inference may be
invalid. You can see how horribly biased the point *and* interval
estimates of every SDT parameter may become when you overly aggregate
your data in the bhsdtr
[preprint](http://dx.doi.org/10.23668/psycharchives.2725). By the way,
I am now in the process of examining the direction and magnitude of
bias in SDT models fitted to overly-aggregated data by using the
excellent Confidence database created by Doby Rahnev.
